---
title: "Tidying Data Notes edX"
author: "Peter Williams"
date: "4/24/2020"
output: pdf_document
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
  
* When tidying data, each row should contain 1 observation and each column should be one variable  

```{r message=FALSE, warning=FALSE, include=FALSE, paged.print=FALSE}
answersE4.2 = c("a", "b", "c", "bcde")
```
  
* Columns can be added with mutate()
  + Glimpse of murder before mutate():

```{r echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}
library(tidyverse)
murders = read.csv("murders.csv")
head(murders, 3)
```
  
  + Glimpse of murder after **mutate()**:
```{r message=FALSE, warning=FALSE, include=FALSE, paged.print=FALSE}
murders = mutate(murders, rate = total / population * 100000)
head(murders, 3)
```
  
* Data sets can be filtered with the **filter()** function:
```{r message=FALSE, warning=FALSE, include=FALSE, paged.print=FALSE}
murders = filter(murders, rate <= 0.71)
head(murders, 3)
```

* Columns can be selected with **select(df, col1, col2, col3, coln)**  
* The **pull()** function can help us to isolate integers from a single observation data frame:
```{r}
us_murder_rate <- murders %>%
  summarize(rate = sum(total) / sum(population) * 100000)
us_murder_rate
# Summarize gives us a new data frame, however, using pull() we can extract integers
us_murder_rate %>% pull(rate)
```
  
* The **arrange()** function can be used to order a dataframe.  
  + if there is a tie in the order, a second argument can be used to break the tie:
  
```{r}
murders %>%
  arrange(region, rate) %>%
  head(2)
# This reads: arrange the murders df by region, if regions are the same arrange the region by rate
```
  
  + Another useful organizational function is the **top_n()** function. This function is the combination of head() and arrange()  


#### Tibbles vs. Data Frames
  
Essentially, a tibble is a modern day data frame, however, there are four major differences between them:  
1.  Tibbles display better  
2.  Subsets of tibbles are tibbles  
3.  Tibbles can have complex entries  
4.  Tibbles can be grouped  
  


